An efficient contact value approach to approximating the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-07-02 papers/pair-correlation/paper.tex
Another round of proof reading.  Made a handful of edits to the text in CVA Efficient section and the A Separable Fit Section Tested on node26. Took 0:41.37.

\section{Making the CVA efficient}
The CVA as described in the previous section is an incremental
improvment over existing approaches outlined in Section~\ref{sec:gn}.
However, as discussed above, these approximations have not been widely
used in the construction of density functionals based on thermodynamic
perturbation theory, largely due to their computational complexity.
We address this issue by developing a \emph{separable} fit to the
radial distribution function of the hard-sphere fluid (see
Section~\ref{sec:separable-fit} for details).  The separable fit is of
the form
\begin{align}
  g_S(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
\end{align}
where the notable aspect is that the radial distribution function is
written as a sum of terms that are each a simple product of a function
of radius with a function of $g_\sigma$.  This enables us to write
typicaly integrals involving the pair distribution function as a
summation of fixed-kernel convolutions, which may be efficiently
computed using Fast Fourier Transforms (FFTs), in $\mathcal{O}(N\log
N)$ time.

@@

We have performed a separable fit to the radial distribution function,
which we compute using Monte Carlo simulation.  We chose to focus our
fit on the range of distances $r_{12} < 2\sigma$.  This range of
distances is relevant to the widely used \fixme{add citations?}
Statistical Associating Fluid Theory of Variable Range (SAFT-VR) free
energy with square-well dispersive attraction developed by
Gil-Villegas \emph{et al.}~\cite{gil1997statistical}.  For comparison,
we plot in Fig.~\ref{fig:radial-distribution} alongside our the fit to
the radial distribution function that of Gil-Villegas \emph{et al.},
which we have extracted from their approximation for the first term in
the dispersion free energy given by Eq.~\ref{eq:mean-field}.

For ease of implementation and future extension to larger radii, we
chose to express our fit using a polynomial fit.  We constrain our
functional form for $g_S(r, g_\sigma)$ to match the value at contact:
\begin{equation}
  g_S(\sigma, g_\sigma) = g_\sigma.
\end{equation}
Beyond this constraint, we simply included a polynomial multiplied by
an exponential decay:
\begin{align}
  g_S(r,g_\sigma) &=
  1 + \left(\sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} h_\sigma^i
  \zeta^j\right)e^{-\alpha \zeta} \label{eq:fit-form}
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function at
contact, and $\zeta \equiv \frac{r}{\sigma}-1$ is a dimensionless
measure of the distance from contact.

@@

The fit is displayed in Fig.~\ref{fig:radial-distribution}, along with
the radial distribution function fit by Gil-Villegas and Monte Carlo
simulation results.  This function is fit to data for $r <= \maxrfit
R$.  The maximum error within this radius in $g_S(r)$ at packing
fractions up to 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$
and $r = \rmaxerr R$.

>>>>

\section{Making the CVA efficient}
The existing approaches to approximating the pair correlation function
outlined in Section~\ref{sec:gn} have not been widely used in the
construction of density functionals based on thermodynamic
perturbation theory, largely due to their computational complexity.
While our CVA as described in the previous section is, in terms of
accuracy, an incremental improvement over these existing approaches,
it makes significant gains in terms of computational efficiency,
allowing for its more managable use in such a construction.  We
address the issue by developing a \emph{separable} fit to the radial
distribution function of the hard-sphere fluid (see
Section~\ref{sec:separable-fit} for details).  This separable fit is
of the form
\begin{align}
  g_S(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
\end{align}
where the notable aspect is that the radial distribution function is
written as a sum of terms that are each a simple product of a function
of radius with a function of $g_\sigma$.  This enables us to write
typicaly \jeffsays{'typical?' maybe just 'integrals'} integrals
involving the pair distribution function as a summation of
fixed-kernel convolutions, which may be efficiently computed using
Fast Fourier Transforms (FFTs), in $\mathcal{O}(N\log N)$ time.

@@

We have performed a separable fit to the Monte-Carlo simulated radial
distribution function.  We chose to focus our fit on the range of
distances $r_{12} < 2\sigma$.  This range of distances is relevant to
the widely used \fixme{add citations?}  Statistical Associating Fluid
Theory of Variable Range (SAFT-VR) free energy with square-well
dispersive attraction developed by Gil-Villegas \emph{et
  al.}~\cite{gil1997statistical}.  For comparison, in
Fig.~\ref{fig:radial-distribution} we plot our fit, Monte-Carlo data,
and the radial distribution function of Gil-Villegas \emph{et al.},
which we have extracted from their approximation for the first term in
the dispersion free energy given by Eq.~\ref{eq:mean-field}.

For ease of implementation and future extension to larger radii, we
chose to express our fit with a polynomial function, multiplied by an
exponential decay.  We constrain our functional form for $g_S(r,
g_\sigma)$ to match the value at contact:
\begin{equation}
  g_S(\sigma, g_\sigma) = g_\sigma.
\end{equation}
Beyond this constraint we have:
\begin{align}
  g_S(r,g_\sigma) &=
  1 + \left(\sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} h_\sigma^i
  \zeta^j\right)e^{-\alpha \zeta} \label{eq:fit-form}
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function at
contact, and $\zeta \equiv \frac{r}{\sigma}-1$ is a dimensionless
measure of the distance from contact.

@@

The Gil-Villegas function is fit to data for $r <= \maxrfit R$.  The
maximum error within this radius in $g_S(r)$ at packing fractions up
to 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr R$.
